---
title: "Bienvenido a BIOL 3740"
author: "BIOMETRIA"
output:
  html_document: 
    theme: spacelab
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```


**Instructor:** Raymond L. Tremblay   
  
**Oficina**: NL 104			 		 
**Teléfono**: (787) 850-0000 (dept de biología)			  		
**Coreo electrónico**: raymond.tremblay_at_upr_dot_edu	
 

### Horario de clase
**Presentación de temas y discución**: M, J at 9:00-10:20am (80 mins) (NOTE: Necesita traer su laptop!)   

**Hora de Estudiantes**:    
- M y J from 1:00 a 3:00pm (NL 104)    


### Libros       
- [Garrett Grolemund, Hadley Wickham. R for Data Science](https://r4ds.had.co.nz) En Ingles. 
- [Garrett Grolemund, Hadley Wickham. R para Ciencia de Datos](https://es.r4ds.hadley.nz) En Español.
- [Tremblay y Hernández-Serrano, 2018](https://www.amazon.com/Gráficas-Versátiles-ggplot2-Análisis-Spanish-ebook/dp/B07P2Q87H2/ref=sr_1_1?dchild=1&keywords=Graficas+versatiles&qid=1591404029&sr=8-1).

•	Articulos revisados por pares serán asignados.    

### Programados
- [R- free statistical programming language](https://cran.r-project.org/) 
- [RStudio](https://rstudio.com)
- MS Excel, Numbers o Google Sheet

### Prerequisites    
- BIOL 3011-3013: Biología General y Laboratorio primer semestre 
- BIOL 3012-3014: Biología General y Laboratorio segundo semestre  

### Descripción del curso  
Estudio de las diferentes técnicas estadísticas con aplicación a la Biología. Se enfatizará en la estadística descriptiva, análisis de regresiones y correlaciones, pruebas de hipótesis paramétricas y no paramétricas y análisis de frecuencias y varianza. Además, se utilizarán las computadoras como mecanismos para facilitar y agilizar el cómputo y análisis estadístico.

### Learning outcomes
Objetivos del curso:
Al finalizar el curso el estudiante podrá:

    1. Discutir la importancia de la estadística en los estudios biológicos.
    2. Ordenar datos biológicos en forma lógica.
    3. Diferenciar entre los diferentes tipos de datos biológicos.
    4. Diferenciar los diferentes tipos de escalas biológicas.
    5. Organizar y resumir datos en forma de tablas.
    6. Confeccionar polígonos de frecuencia.
    7. Creación de gráficas para demostrar patrones.
    8. Hacer uso adecuado de las diferentes estadísticas descriptivas.
    9. Describir las diferentes medidas de tendencia central.
    10. Basado en las medidas de tendencia central, inferir algunas de las características de la muestra bajo estudio.
    11. Determinar cuándo se deben calcular en forma grupal las medidas de tendencia central y dispersión de la muestra.
    12. Calcular probabilidades en situaciones específicas.
    13. Establecer la diferencia entre distribuciones discretas y continuas.
    14. Formular una hipótesis nula y alterna, adecuada para una situación en particular.
    15. Distinguir de acuerdo al tipo de datos recolectados cuándo utilizar pruebas paramétricas o pruebas no paramétricas.
    16. Evaluar los supuestos de las pruebas, tal como distribución normal e igualdad de varianza
    17. Evaluar la hipótesis nula y determinar el grado de significación de esta hipótesis haciendo uso de los diferentes análisis estadísticos.
    18. Confeccionar tablas de contingencia y establecer homogeneidad.
    19. Determinar cuándo se debe utilizar un análisis de regresión.
    20. Calcular coeficientes de regresión.
    21. Determinar intervalos de confianza y significación de una regresión.
    22. Calcular el coeficiente de correlación.
    23. Realizar comparaciones entre coeficientes de correlación simples.
    24. Desarrollar y analizar diseños experimentales utilizando técnicas de muestreo y recolección de datos apropiados.
    25. Determinar el error de muestra.
    26. Conocer y aplicar las pruebas básicas en estadísticas.
    27. Evaluar situaciones experimentales y poder desarrollar un diseño experimental adecuado y lógico para analizar el problema y llegar a conclusiones.
    28. Se hará énfasis en el uso de programados estadísticos para calcular los índices y probar hipótesis. Los programa de computadora puede incluir R, SPSS, Minitab o otros.  

### Grading: 
The course grade will be based on the following components:

Lab exercises (8 total)	         20%  (80 points)     
Quizzes and participation        10%  (40 points)     
Group project                    25%  (100 points)      
Midterm exam # 1 (date TBD)	     10%  (40 points)    
Midterm exam # 2 (date TBD)   	 10%  (40 points)    
Final exam (5/11/2018)   	       25%  (100 points)    

NOTE: Graduate students enrolled in NRES 670 will have an additional 50 pts used to calculate their grade (see below) of a total of 370 points. 
Grading scale: A (100 to 93), A- (92 to 90), B+ (89 to 87), B (86 to 83), B- (82 to 80), C+ (79 to 77), C (76 to 73), C- (72 to 70), D+ (69 to 67), D (66 to 63), D- (62 to 60), F (below 60). 

### Exams:
There will be two midterm exams and a comprehensive final exam. These will consist of multiple-choice, short-answer questions, and essay questions requiring synthesis of ideas and critical thinking. The midterm exam will be cumulative, and based on all information presented up through the week prior to the exam.

### Lectures
Lecture grades will be based primarily on participation and occasional short quizzes. Participation is essential to the learning process (and to our mutual enjoyment of this class). Learning is not a passive process; students are expected to engage with the material in class rather than simply listen and take notes. You should be prepared in class to ask questions, to answer questions posed by other students, and to engage in frequent problem-solving activities (in class). 

### Labs  
Lab exercises will focus on applying concepts and methods introduced in lectures, and will involve real data and problems in wildlife conservation and management wherever possible. Graded lab assignments will involve figures, tables, InsightMaker models and R code (when applicable) and responses to questions in short-answer format. Laboratory write-ups will be due the following lab period, unless otherwise specified. 

### Final group project 
Students will work in groups of ~2-3 to perform a population viability analysis (PVA) to rank conservation or management actions for a species of conservation concern (species of your choice!). Grading will be based on finished products (written and oral presentations) as well as participation and peer evaluations. 

### Graduate credit (for students enrolled in NRES 670) 
Graduate students will be subject to additional expectations in order to receive graduate credit for this course. In particular, graduate students will be expected to develop an original lecture and lead an original lab activity. Graduate students will also be expected to achieve a deeper understanding of the course material, and therefore will be assigned additional readings from the scientific literature and will be expected to participate as leaders in discussions and lab activities.

### Make-up policy and late work:
Missed exams and labs cannot be made up, except in the case of emergencies. If you miss a class meeting, it is your responsibility to talk to one of your classmates about what you missed. If you miss a lab meeting, you are still responsible for completing the lab activities and write-up on your own time. You do not need to let me know in advance that you are going to miss class or lab.

### Students with Disabilities
Any student with a disability needing academic adjustments or accommodations is requested to speak with the Disability Resource Center (Thompson Building, Suite 101) as soon as possible to arrange for appropriate accommodations.

### Statement on Academic Dishonesty
Cheating, plagiarism or otherwise obtaining grades under false pretenses constitute academic dishonesty according to the code of this university. Plagiarism is using the ideas or words of another person without giving credit to the original source; this includes copying another student in class. Always cite the source of your information. This includes copying or paraphrasing from a book, journal, or unpublished material without giving credit to the author(s), and submitting a term paper that was used in another course. Academic dishonesty will not be tolerated and penalties can include filing a final grade of "F"; reducing the student's final course grade one or two full grade points; awarding a failing mark on the coursework in question; or requiring the student to retake or resubmit the coursework. For more details, see the [University of Nevada, Reno General Catalog](http://catalog.unr.edu/).

### Statement on Audio and Video Recording
Surreptitious or covert video-taping of class or unauthorized audio recording of class is prohibited by law and by Board of Regents policy. This class may be videotaped or audio recorded only with the written permission of the instructor. In order to accommodate students with disabilities, some students may have been given permission to record class lectures and discussions. Therefore, students should understand that their comments during class may be recorded.

### Statement for Academic Success Services
Your student fees cover usage of the University Math Center [(775) 784-4433], University Tutoring Center [(775) 784-6801], and [University Writing Center (775) 784-6030]. These centers support your classroom learning; it is your responsibility to take advantage of their services.

### This is a safe space
The University of Nevada, Reno is committed to providing a safe learning and work environment for all. If you believe you have experienced discrimination, sexual harassment, sexual assault, domestic/dating violence, or stalking, whether on or off campus, or need information related to immigration concerns, please contact the University’s Equal Opportunity & Title IX Office at 775-784-1547. Resources and interim measures are available to assist you. For more information, please visit: http://www.unr.edu/equal-opportunity-title-ix .”

### Class protocol
All mobile electronic devices are to be turned off during class unless the instructor gives advance permission (but laptop computers will be used in class regularly).


```{r echo=FALSE, eval=FALSE}
rmarkdown::render('index.Rmd', 'word_document')
rmarkdown::render('schedule.Rmd', 'word_document')
rmarkdown::render('LAB1.Rmd', 'word_document')
rmarkdown::render('LAB2.Rmd', 'word_document')
rmarkdown::render('LAB3.Rmd', 'word_document')
rmarkdown::render('LAB4.Rmd', 'word_document')
rmarkdown::render('LAB5.Rmd', 'word_document')
rmarkdown::render('LAB6.Rmd', 'word_document')
rmarkdown::render('LAB7.Rmd', 'word_document')
## [Quiz #1](https://unr.instructure.com/courses/15896/quizzes/21572)
## [Quiz #2](https://unr.instructure.com/courses/15896/quizzes/24149)
### Field Trips
# We will be taking 1-2 field trips to collect wildlife population data. One field trip will be full-day, on a weekend: bring breakfast, lunch, and plenty of water. Bring any appropriate gear for spending the day in the field including: binoculars, spotting scopes, and field guides, and wear appropriate clothing for hiking. Be prepared for inclement weather.
## UPCOMING
- metapopulations
- source sink
```